---
title: "Diversity effects: bird community"
author: "collnell"
date: "July 5, 2016"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: tango
    fig_caption: true
---

```{r setup, include=FALSE,echo=F}
library(shiny)
library(ggplot2)
library(cowplot)
library(reshape2)
library(plotly)
library(dplyr)
library(labdsv)
library(knitr)
library(pander)
library(rich)
library(car)
###enter data
birds<-read.csv("data/enemy_bird_visit.csv")
sp.list<-read.csv("data/enemy_bird_list.csv")
plots<-read.csv("data/enemy_plots.csv")
```

```{r,datawrangling,warning=F,include=F,echo=F}
std <- function(x) sd(x)/sqrt(length(x))

##clean data 
drops<-c("Grand.Total","X.blank.")##clean up df
birds=birds[,!(names(birds) %in% drops)]
birds[is.na(birds)]<-0
cols<-names(birds)
drops<-c("DIVERSITY","PLOT","VISIT")
sps<-cols[-c(1:3)]##this is a list of the names of the birds only
birds$VISIT<-as.numeric(birds$VISIT)
birds$total_abun<-rowSums(birds[,sps])
bird.sp<-birds[,sps]#matrix plot x sp
plots[is.na(plots)]<-0
plots$PLOT<-as.character(plots$PLOT)
plots$CEIBA<-as.factor(plots$CEIBA)

###diversity matrices
bird.poly<-birds%>%
  filter(DIVERSITY == 'P')
bird.poly.mat<-bird.poly[,sps]
bird.mono<-birds%>%
  filter(DIVERSITY == 'M')
bird.mono.mat<-bird.mono[,sps]
#rarefaction/sp acuum
rarc.poly<-rarc(bird.poly.mat,nrandom=99)
rarc.mono<-rarc(bird.mono.mat,nrandom=99)
```

```{r,datasummary,echo=F,warning=F}
##create dfs summarizing at visit, plot, and diversity levels forbirds ocmmunity metrics
###by plot visit
birds$total_abun<-rowSums(birds[,sps])
birds$shan<-diversity(bird.sp)##plot/visit level shannon calculation
birds$richness<-specnumber(bird.sp)##no adjustment for indivudals

##by plot
birds.plot.mean.vars<-birds%>%##plot level means (per visit)
  group_by(DIVERSITY,PLOT)%>%
  summarize(n_visit = max(VISIT),
            mean_abun = mean(total_abun), se_abun = std(total_abun), total_abun = sum(total_abun),
            mean_shan = mean(shan), se_shan = std(shan),
            mean_rich = mean(richness), se_rich = std(richness))
birds.plot.mean.vars$PLOT<-as.character(birds.plot.mean.vars$PLOT)##missing total richness, total diversity based on plot sums
birds.plot.sum.vars<-birds%>%
  group_by(DIVERSITY,PLOT)%>%
  summarize_each(funs(sum))
plot.sp<-birds.plot.sum.vars[,sps]
birds.plot.sum.vars$plot_shan<-diversity(plot.sp)
birds.plot.sum.vars$plot_rich<-specnumber(plot.sp)
birds.plot.sum.vars$PLOT<-as.character(birds.plot.sum.vars$PLOT)##includes summed plot shannon diversity, abundance, and sp richness
###single plot-level df
birds.plot.vars<-left_join(birds.plot.mean.vars,birds.plot.sum.vars[,c("PLOT","plot_shan","plot_rich")],by="PLOT")

#by plot diversity
birds.div.mean.vars<-birds.plot.mean.vars%>% ##diversity level means (per visit)
  group_by(DIVERSITY)%>%
  summarize(total_visit_div = sum(n_visit),n_plots_div = length(PLOT),
            total_abun_div = sum(total_abun),mean_abun_div = mean(mean_abun), se_abun_div = std(mean_abun),
            mean_shan_div = mean(mean_shan), se_shan_div = std(mean_shan),
            mean_rich_div = mean(mean_rich), se_rich_div = std(mean_rich))##missing diversity summed values
birds.plot.sum.vars$PLOT<-as.numeric(birds.plot.sum.vars$PLOT)
birds.div.sum.vars<-birds.plot.sum.vars%>%
  group_by(DIVERSITY)%>%
  summarize_each(funs(sum))
div.sp<-birds.div.sum.vars[,sps]  
birds.div.sum.vars$div_shan<-diversity(div.sp)
birds.div.sum.vars$div_rich<-specnumber(div.sp)  
birds.div.vars<-left_join(birds.div.mean.vars,birds.div.sum.vars[,c("DIVERSITY","div_shan","div_rich")],by="DIVERSITY")

````

```{r,treeID,echo=F,warning=F}
birds.plot.vars.trees<-left_join(birds.plot.vars,plots[,c("PLOT","GENODIV","CAOBA","CEIBA","CIRI","JABIN","MACU","PICH")],by="PLOT")
```

```{r, melteddf,echo=F,message=F,warning=F}
birds.melt<-melt(birds,id.vars=c("DIVERSITY","PLOT","VISIT","total_abun"),variable.name="ID")%>%
  left_join(sp.list[,c("ID","feeding.guild","Family","Order")],by="ID") ##merge with taxonomic groupings
```

```{r,insectivores,echo=F,message=F,warning=F}
insect.birds<-birds.melt%>% ##filter data to insectivorous birds
  filter(feeding.guild %in% c("IN","FL","OM"))

ins.plot<-insect.birds%>%
  group_by(DIVERSITY,PLOT,VISIT)%>%
  summarize(total_birds_plot_visit=sum(value))

ins.plot1<-ins.plot%>%
  group_by(DIVERSITY,PLOT)%>%
  summarize(total_birds_plot = sum(total_birds_plot_visit),
            mean_birds_visit = mean(total_birds_plot_visit),
            n_visit=as.numeric(max(VISIT)),
            se_birds_visit = std(total_birds_plot_visit))

```  

```{r,plotting,echo=F}
plot_means<-function(){
  list(
    stat_summary(fun.y = "mean", geom = "bar", fill = DIVERSITY),
    stat_summary(fun.data = "mean_cl_normal", geom = "errorbar", width = 0.2)
  )
}

plot_aes<-function(){
  list(
    scale_color_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77")),
    theme_minimal()
  )
}
```
These data are from a field experiment conducted July 2015 at the [UADY forest diversity plots near Muna, Yucatan, Mexico](http://www.treedivnet.ugent.be/ExpUADY.html). Using the tree diversity manipulation we test the predictions that  
(1) tree diversity promotes the abundance and diversity of predators (insectivorous birds) and   
(2) via effects on predators, tree diversity is correlated with greater herbivore suppression.  
  
#Bird Observations: All Birds  
###Abundance  
Does **bird abundance** differ between **monoculture** and **polyculture** plots?  
```{r, fguilds,fig.height=4,message=F,echo=F,warning=F}
##summarizing data by feeding guilds at multiple scales

#plot diversity
div.guild<-birds.melt%>%###this is the total number of birds~diversity+feeding guild
  group_by(DIVERSITY,feeding.guild)%>%
  summarize(total_birds = sum(value))

##visit
plot.visit<-birds.melt%>%
  group_by(DIVERSITY,PLOT,VISIT)%>%
  summarize(total_birds_plot=sum(value))%>%
  group_by(DIVERSITY,PLOT)%>%
  summarize(mean_birds_plot = mean(total_birds_plot))

##plot
plot.guild<-birds.melt%>%
  group_by(DIVERSITY,PLOT,VISIT,feeding.guild)%>%
  summarize(total_birds_visit = sum(value))%>%
  group_by(DIVERSITY,PLOT,feeding.guild)%>%
  summarize(mean_birds_visit=mean(total_birds_visit),se_birds_visit = std(total_birds_visit))%>%
  left_join(plot.visit[,c("PLOT","mean_birds_plot")],by="PLOT")%>%
  mutate(per_birds_visit= (mean_birds_visit/mean_birds_plot)*100)

plot.dcast<-dcast(plot.guild,DIVERSITY+PLOT~feeding.guild)##div + plot ~guilds in percentages

plot.guild2<-birds.melt%>%
  group_by(DIVERSITY,PLOT,VISIT,feeding.guild)%>%
  summarize(total_birds_visit = sum(value))

plot.guild3<-plot.guild2%>%
  group_by(DIVERSITY,PLOT,feeding.guild)%>%
  summarize(mean_birds_visit=mean(total_birds_visit),se_birds_visit = std(total_birds_visit))%>%
  group_by(DIVERSITY,feeding.guild)%>%
  summarize(mean_birds_div = mean(mean_birds_visit), se_birds_div = std(mean_birds_visit))

```
<br>  
```{r,abundance.raw,fig.height=4,echo=F}
##histogram of raw data
abun.hist<-ggplot(birds,aes(x=total_abun,fill=DIVERSITY,color=DIVERSITY),alpha=.6)+geom_density(alpha=.7)+
  theme_minimal()+labs(y="Density",x="Bird Abundance")+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+plot_aes()+
  theme(legend.position="top")
##raw observations
abun.rarc.ind<-ggplot()+#by individuals
  geom_point(data=rarc.poly,aes(x=samples,y=individuals),size=1,shape=1,color="#1B9E77")+
  geom_point(data=rarc.mono,aes(x=samples,y=individuals),size=1,shape=1,color="darkslateblue")+
  theme_minimal()+labs(y="Individuals",x="Observation")

plot_grid(abun.hist,abun.rarc.ind)
```


```{r,abundance.plot,fig.height=3.5,message=F,echo=F}

abun.plots<-ggplot(birds.plot.vars,aes(x=reorder(PLOT,mean_abun),y=mean_abun,color=DIVERSITY))+
  geom_errorbar(aes(ymin=mean_abun-se_abun,ymax=mean_abun+se_abun),color="black")+
  geom_point(size=3,shape=16)+
  scale_color_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+
  theme_minimal()+
  theme(legend.position="none",axis.text.x=element_blank())+
  labs(x="Plot",y="Birds/Visit")
abun.plot.mean<-abun.plots+geom_point(data=birds,aes(x=reorder(PLOT,total_abun),y=total_abun),size=1,shape=1)


abun.div<-ggplot(birds.div.vars,aes(x=DIVERSITY,y=mean_abun_div,color=DIVERSITY))+
  geom_point(size=8,shape=19)+
  geom_errorbar(aes(ymin=mean_abun_div-se_abun_div,ymax=mean_abun_div+se_abun_div),color="black",width=.1)+
  labs(x="Plot Diversity",y="Birds/Visit")+
  theme_minimal()+
  scale_color_manual(labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+theme(legend.position="none")
abun.div.mean<-abun.div+geom_point(data=birds.plot.vars,aes(x=DIVERSITY,y=mean_abun),size=3,shape=1,alpha=.65)

plot_grid(abun.plot.mean,abun.div.mean)
```
  
  
ANOVA testing for main effect of tree DIVERSITY controlling for PLOT    
```{r,abuntests, results="asis",echo=F}
birds$PLOT<-as.character(birds$PLOT)
birds$VISIT<-as.character(birds$VISIT)
abun.by.div<-aov(total_abun~DIVERSITY+PLOT,data=birds)
pander(summary(abun.by.div),round=c(0,2,2,2,3))  
```
<style type="text/css">
.table {
    width: 50%;
}
</style>  
<br> 
<br>   
and among **feeding guilds?**

```{r,abundance.fg,fig.height=3.5,echo=F}

fg.abun<-ggplot(div.guild,aes(y=total_birds,x=reorder(feeding.guild,total_birds),fill=DIVERSITY))+
  geom_bar(stat="identity")+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+
  theme_minimal()+theme(legend.position="top")+labs(x="Feeding Guild",y="Total Birds Observed")

fg.plot.abun<-ggplot(plot.guild3,aes(x=reorder(feeding.guild,mean_birds_div),y=mean_birds_div,fill=DIVERSITY))+
  geom_bar(stat="identity",position="dodge")+theme_minimal()+scale_fill_manual(values=c("darkslateblue","#1B9E77"))+
  geom_errorbar(aes(ymin=mean_birds_div-se_birds_div,ymax=mean_birds_div+se_birds_div),width=.2,position=position_dodge(.9))+theme(legend.position="none")+
  labs(x="Feeding Guild",y="Birds/Visit")

plot_grid(fg.abun,fg.plot.abun)
```  

Guilds: Granivorous (GR), Carnivorous (CA), Nectivorous (NE), Frugivorous (FR), Insectivorous (FL = flycatching, IN= gleaning), Omnivorous (OM), NA = unidentified   
  
  
    
ANOVA testing for main and interactive effects of tree DIVERSITY and FEEDING GUILD
```{r,abun.fg.tests, results="asis",echo=F}    
abun.by.fg<-lm(total_birds_visit~DIVERSITY*feeding.guild+PLOT,data=plot.guild2)
pander(summary(aov(abun.by.fg)),round=c(0,2,2,2,3)) 
```
 <style type="text/css">
.table {
    width: 50%;
}
</style>  
<br> 
<br>


###Richness  
 
  
```{r,richness,include=F,echo=F}
allrichness<-rich(matrix=bird.sp,verbose=T)
##dont include hard to get rid of message
```

```{r,richness.fg,echo=F}

birds.div.vars<-birds.plot.mean.vars%>%
  group_by(DIVERSITY)%>%
  summarize(total_abun = sum(total_abun),mean_abun_div = mean(mean_abun), se_abun_div = std(mean_abun),mean_shan_div = mean(mean_shan), se_shan_div = std(mean_shan), mean_rich_div = mean(mean_rich), se_rich_div = std(mean_rich))

fgrich<-birds.melt%>%
  group_by(DIVERSITY,feeding.guild,ID)%>%
  summarize(guild.rich = sum(value))
fg2<-fgrich%>%
  group_by(DIVERSITY,feeding.guild)%>%
  summarize(richn = sum(guild.rich>0))
##these are diversity-level means % se based on plot/visit means
```
Mean richness plot/visit: `r allrichness$mr`  
Total richness: `r allrichness$cr`  
Singletons: `r allrichness$singletons`  


```{r,richness.rarc,fig.height=4,message=F,comment=F,echo=F}

div.rarc<-ggplot()+#by samples
  geom_point(data=rarc.poly,aes(x=samples,y=richness),size=1,shape=1,color="#1B9E77")+
  geom_point(data=rarc.mono,aes(x=samples,y=richness),size=1,shape=1,color="darkslateblue")+
  theme_minimal()+labs(x="Observation",y="Species Richness")+scale_color_manual("Plot Diversity",labels=c("Monoculture","Polyculture"))+
  theme(legend.position="top")

div.rarc.ind<-ggplot()+#by individuals
  geom_point(data=rarc.poly,aes(x=individuals,y=richness),size=1,shape=16,color="#1B9E77")+
  geom_point(data=rarc.mono,aes(x=individuals,y=richness),size=1,shape=16,color="darkslateblue")+
  theme_minimal()+labs(x="Individuals",y="Species Richness")

plot_grid(div.rarc,div.rarc.ind)

```  
  
  
  
**Richness estimators:**    
```{r,richness.est,results="asis",echo=F}
all.spec<-specpool(bird.sp)
poly.spec<-specpool(bird.poly.mat)
mono.spec<-specpool(bird.mono.mat)
spec.res<-rbind(all.spec,poly.spec,mono.spec)
rownames(spec.res)<-c("All Plots","Polycultures","Monocultures")
colnames(spec.res)<-c("Observed","Chao2","Chao2_se","Jack1","Jack1_se","Jack2","Boot","Boot_se","N")
pander(spec.res[,c(1,2,4,7,9)]) 
```
Incidence-based extrapolation to estimate **species richness** using 'specpool' in 'vegan'.
  <style type="text/css">
.table {
    width: 80%;
}
</style>  
<br>  
```{r,richness.raw,fig.height=4,echo=F}
#histogram
hist.rich<-ggplot(birds,aes(x=richness,fill=DIVERSITY,color=DIVERSITY),alpha=.6)+geom_density(alpha=.7)+
  theme_minimal()+labs(y="Density",x="Species Richness")+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+plot_aes()+
  theme(legend.position="top")##histogram of richness values by diversity

rich.div<-ggplot(data=birds.div.vars,aes(x=DIVERSITY,y=mean_rich_div,color=DIVERSITY))+##richness plot by diversity
  geom_point(size=8,shape=19)+
  geom_errorbar(aes(ymin=mean_rich_div-se_rich_div,ymax=mean_rich_div+se_rich_div),color="black",width=.1)+
  labs(x="Plot Diversity",y="Mean Richness/Visit")+
  theme_minimal()+
  scale_color_manual(labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+theme(legend.position="none")
rich.div.mean<-rich.div+geom_point(data=birds.plot.vars,aes(x=DIVERSITY,y=mean_rich),size=3,shape=1,alpha=.65)

plot_grid(hist.rich,rich.div.mean)
```  
  
ANOVA testing for main effect of tree DIVERSITY & PLOT    
```{r,rich.test, results="asis",echo=F}
rich.by.div<-aov(richness~DIVERSITY+PLOT,data=birds)
pander(summary(rich.by.div))  
```
<style type="text/css">
.table {
    width: 50%;
}
</style>  
<br>  
**Species richness** by **feeding guild**
```{r,richguild,fig.height=4,echo=F}
fg_me<-birds.melt%>%
  group_by(DIVERSITY,PLOT,VISIT,feeding.guild)%>%
  summarize(visit_rich= sum(value>0))
fg_mean<-fg_me%>%
  group_by(DIVERSITY,PLOT,feeding.guild)%>%
  summarize(plot_rich = mean(visit_rich))%>%
  group_by(DIVERSITY,feeding.guild)%>%
  summarize(div_rich = mean(plot_rich), se_div_rich = std(plot_rich))##mean plkot richness

guildrich<-ggplot(fg2,aes(y=richn,x=reorder(feeding.guild,richn),fill=DIVERSITY))+
  geom_bar(stat="identity")+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+
  theme_minimal()+theme(legend.position="top")+labs(x="Feeding Guild",y="Total Species Richness")
divguild.rich<-ggplot(fg_mean,aes(x=reorder(feeding.guild,div_rich),y=div_rich,fill=DIVERSITY))+
  geom_bar(stat="identity",position="dodge")+theme_minimal()+scale_fill_manual(values=c("darkslateblue","#1B9E77"))+
  geom_errorbar(aes(ymin=div_rich-se_div_rich,ymax=div_rich+se_div_rich),width=.2,position=position_dodge(.9))+theme(legend.position="none")+
  labs(x="Feeding Guild",y="Mean Richness/Visit")

plot_grid(guildrich,divguild.rich)
```  
  
ANOVA testing for effects of tree DIVERSITY and FEEDING GUILD
```{r,richbydivfg, results="asis",echo=F}    
rtc<-lm(visit_rich~DIVERSITY*feeding.guild+PLOT,data=fg_me)
pander(summary(aov(rtc)),round=c(0,2,2,2,3)) 
```
 <style type="text/css">
.table {
    width: 50%;
}
</style>    
  
**Shared species** 

###Diversity  
Shannon Diversity  
```{r,divhist,fig.height=4,echo=F}

shan.hist<-ggplot(birds,aes(x=shan,fill=DIVERSITY,color=DIVERSITY),alpha=.6)+geom_density(alpha=.7)+
  theme_minimal()+labs(y="Density",x="Shannon Diversity")+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+plot_aes()+
  theme(legend.position="top")##histogram of div values by diversity

shan.div<-ggplot(data=birds.div.vars,aes(x=DIVERSITY,y=mean_shan_div,color=DIVERSITY))+##rshannn by diversity
  geom_point(size=8,shape=19)+
  geom_errorbar(aes(ymin=mean_shan_div-se_shan_div,ymax=mean_shan_div+se_shan_div),color="black",width=.1)+
  labs(x="Plot Diversity",y="Mean Diversity/Visit")+
  theme_minimal()+
  scale_color_manual(labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+theme(legend.position="none")
shan.div.mean<-shan.div+geom_point(data=birds.plot.vars,aes(x=DIVERSITY,y=mean_shan),size=3,shape=1,alpha=.65)

plot_grid(shan.hist,shan.div.mean)
```  
  
ANOVA testing for main effect of tree DIVERSITY & PLOT    
```{r,shan.tests, results="asis",echo=F}
shan.by.div<-aov(shan~DIVERSITY+PLOT,data=birds)
pander(summary(shan.by.div))  
```
<style type="text/css">
.table {
    width: 50%;
}
</style>  
<br>  
  

###Phylogenetic Diversity   

###Species Composition  


###Taxonomy  

**`r length(unique(birds.melt$PLOT))` plots** visited 4x each   
**`r sum(birds.melt$value)` bird observations** of **`r length(unique(birds.melt$ID))` species** from **`r length(unique(birds.melt$Family))` families** (`r sort(unique(birds.melt$Family))`) and **`r length(unique(birds.melt$Order))` orders** (`r sort(unique(birds.melt$Order))`)  
 
#Tree Identity Effects  
```{r,treeIDfx,echo=F,message=F,fig.height=4}
birds.plot.vars.trees<-left_join(birds.plot.vars,plots[,c("PLOT","GENODIV","CAOBA","CEIBA","CIRI","JABIN","MACU","PICH")],by="PLOT")
trees<-birds.plot.vars.trees[,c("DIVERSITY","PLOT","mean_abun","mean_shan","mean_rich","CAOBA","CEIBA","CIRI","JABIN","MACU","PICH")]
tree.melt<-melt(trees,id.vars=c("DIVERSITY","PLOT","CAOBA","CEIBA","CIRI","JABIN","MACU","PICH"))
tree.melt$CEIBA<-as.factor(tree.melt$CEIBA)
tree.melt$CAOBA<-as.factor(tree.melt$CAOBA)
tree.melt$CIRI<-as.factor(tree.melt$CIRI)
tree.melt$JABIN<-as.factor(tree.melt$JABIN)
tree.melt$MACU<-as.factor(tree.melt$MACU)
tree.melt$PICH<-as.factor(tree.melt$PICH)
##for tests
abundf<-filter(tree.melt,variable=="mean_abun")
richdf<-filter(tree.melt,variable=="mean_rich")
shandf<-filter(tree.melt,variable=="mean_shan")
abun_melt<-melt(abundf,id.vars=c("DIVERSITY","PLOT","variable","value"),variable.name="TREE",value.name="sp")
abun.tree<-abun_melt%>%
  group_by(DIVERSITY,TREE,sp)%>%
  summarize(mean_abun_sp = mean(value),se_abun_sp = std(value),sd_abun_sp = sd(value),n_plots = length(PLOT))
levels(abun.tree$DIVERSITY)<-c("Monoculture","Polyculture")
abun.tree$sp<-as.factor(abun.tree$sp)
levels(abun.tree$sp)<-c("with","without")
trcast<-dcast(abun.tree,DIVERSITY+TREE~sp,value.var=c("mean_abun_sp"))##2 cols for sp
trcastsd<-dcast(abun.tree,DIVERSITY+TREE~sp,value.var=c("sd_abun_sp"))%>%
  rename(with_sd = with, without_sd = without)##2 cols for sp
trcastn<-dcast(abun.tree,DIVERSITY+TREE~sp,value.var=c("n_plots"))%>%
  rename(with_n = with, without_n = without)##2 cols for sp
trcast$fx<-log((trcast$without)/(trcast$with))##proportional difference in number of species between control & treatment (with)
#ex=C=without
fxdf<-cbind(trcast[,c("DIVERSITY","TREE","with","without","fx")],trcastsd[,c("with_sd","without_sd")],trcastn[,c("with_n","without_n")])
#ex=C=without

#df$se<-(sqrt((df$C_sd)^2/(df$C_n*df$C_mean)+(df$T_sd)^2/(df$T_n*df$T_x)))/(df$T_n+df$C_n)
fxdf$fx_se<-(sqrt((fxdf$without_sd)^2/(fxdf$without_n*fxdf$without)+(fxdf$with_sd)^2/(fxdf$with_n*fxdf$with)))/(fxdf$with_n+fxdf$without_n)

```

```{r,treeIDfxplots,echo=F,message=F,fig.height=4,warning=F}
tree.abun.div4<-ggplot(data=abun.tree,aes(x=TREE,y=mean_abun_sp,color=DIVERSITY))+
  geom_errorbar(aes(ymin=mean_abun_sp-se_abun_sp,ymax=mean_abun_sp+se_abun_sp),width=.1,color="black")+
  geom_point(data=abun.tree,aes(shape=sp),size=3)+
  scale_shape_manual("Species Present",labels=c("No","Yes"),values=c(1,16))+
  scale_color_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+
  facet_grid(.~DIVERSITY)+labs(x="Tree Species",y="Bird Observations/Visit")+
    theme_minimal()+theme(axis.text.x=element_text(angle=45))+guides(color="none")
tree.abun.div4 ##abundance by div~tree+presense



tree.abun.div5<-ggplot(data=fxdf,aes(x=TREE,y=fx,fill=DIVERSITY))+
  geom_bar(data=fxdf,stat="identity")+
  geom_errorbar(aes(ymin=fx-fx_se,ymax=fx+fx_se),width=.2,color="black",position=position_dodge(.9))+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+
  facet_grid(.~DIVERSITY)+labs(x="Tree Species",y="Effect Size (Bird Abundance)")+
  theme_minimal()+theme(axis.text.x=element_text(angle=45))+
  geom_hline(aes(yintercept=0),lty="dashed")
tree.abun.div5 ##ab
##effect size by species/diversity with errorbars
tree.abun.div6<-ggplot(data=fxdf,aes(x=TREE,y=fx,fill=DIVERSITY))+
  geom_bar(data=fxdf,stat="identity",position="dodge")+
  geom_errorbar(aes(ymin=fx-fx_se,ymax=fx+fx_se),width=.2,color="black",position=position_dodge(.9))+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+
  labs(x="Tree Species",y="Effect Size (Bird Abundance)")+
  theme_minimal()+theme(axis.text.x=element_text(angle=45))+
  geom_hline(aes(yintercept=0),lty="dashed")
tree.abun.div6 ##ab
```


#Bird Obs: Insectivorous Birds   
(including gleaners, flycatchers, and omnivores that consume arthropods)  
###same with filtered data
```{r, insguilds,fig.height=4.5,echo=F,warning=F}
ins.guild<-insect.birds%>%###this is the total number of insecitvorous birds~diversity+feeding guild
  group_by(DIVERSITY,feeding.guild)%>%
  summarize(total_birds = sum(value))

guilddiv<-ggplot(ins.guild,aes(y=total_birds,x=reorder(feeding.guild,total_birds),fill=DIVERSITY))+
  geom_bar(stat="identity")+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+
  theme_minimal()+theme(legend.position="top")+labs(x="Feeding Guild",y="Total Birds Observed")

abun_by_plot<-ggplot(ins.plot1,aes(x=reorder(PLOT,mean_birds_visit),y=mean_birds_visit,fill=DIVERSITY))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=mean_birds_visit-se_birds_visit,ymax=mean_birds_visit+se_birds_visit))+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=c("darkslateblue","#1B9E77"))+
  theme_minimal()+
  theme(legend.position="none",axis.text.x=element_blank())+
  labs(x="Plot",y="Birds Observed/Visit")

plot_grid(guilddiv,abun_by_plot)
```


**`r sum(insect.birds$value)` observations** of **`r length(unique(insect.birds$ID))` species** from **`r length(unique(insect.birds$Family))` families** (`r sort(unique(insect.birds$Family))`) and **`r length(unique(insect.birds$Order))` orders**(`r sort(unique(insect.birds$Order))`)  
  
```{r,insbytotal,fig.height=3.5,fig.width=3.5,echo=F}
birds$PLOT<-as.character(birds$PLOT)
ins.plot$PLOT<-as.character(ins.plot$PLOT)
compareins<-left_join(birds,ins.plot,by="PLOT")
ins.by.total<-lm(total_birds_plot_visit~total_abun,data=compareins)
p<-summary(ins.by.total)$coefficients[,4]
r2<-summary(ins.by.total)$r.squared

typeplot<-ggplot(data=compareins,aes(x=total_abun,y=total_birds_plot_visit))+
  geom_jitter(data=compareins,aes(x=total_abun,y=total_birds_plot_visit,color=DIVERSITY.x),size=1,shape=1)+
  geom_smooth(method="lm",se=F,color="black")+
  theme_minimal()+
  scale_color_manual(values=c("darkslateblue","#1B9E77"))+
  theme(legend.position="none")+
  labs(x="Birds/Visit",y="Insectivores/Visit")
typeplot

```
    
  
#Predation Experiment  

#Data  
###Description
These data are from a field experiment conducted July 2015 at the [UADY forest diversity plots near Muna, Yucatan, Mexico](http://www.treedivnet.ugent.be/ExpUADY.html). Using the tree diversity manipulation we test the predictions that  
(1) tree diversity promotes the abundance and diversity of predators (insectivorous birds) and   
(2) via effects on predators, tree diversity is correlated with greater herbivore suppression.  
  
###Data sheets  
[**Bird Observations**](https://github.com/collnell/Enemies/blob/master/enemy_bird_visit.csv)  
Bird survey results from 15 minute point counts at the plot level. Each plot was visited 4 times on separate days between 0600 and 0930. These observations reflect birds actively foraging in the plots, excluding birds flying overhead or through the plot. The species of trees used by foraging birds were recorded once per individual tree. Birds that were not able to be identified definitively are recorded as species "X", or "CH X" if a partial ID was known (CH X = chipe unknown (warbler). All bird IDs match species information on the bird list.     
  
[**Bird List**](https://github.com/collnell/Enemies/blob/master/enemy_bird_list.csv)   
Inventory on all bird species recorded at the UADY diversity plots in January and July of 2015.  Contians species codes and associated information for each ID (including feeding guild).  

**Predation Experiment**  
These data are from a clay caterpillar experiment used to assay predation rates between the plots. For each experimental plot, 2 clay models were placed on opposite branches of 12 trees. Models were visited twice, 24 and 48 hours after placement, to record attack marks. Marked models were identified as either Arthropod, Bird, and Mammal and replced with a new model.  

**Experimental Design**  
Data on plots used in experiment. Plot diversity, tree species, ETC  


