
---
title: "Diversity effects: the enemies hypothesis"
author: "collnell"
date: "July 5, 2016"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: tango
    fig_caption: true
    code_folding: hide
---

```{r setup, include=FALSE}
library(shiny)
library(ggplot2)
library(cowplot)
library(reshape2)
library(plotly)
library(dplyr)
library(labdsv)
library(wesanderson)
library(knitr)
library(pander)
###enter data
birds<-read.csv("enemy_bird_visit.csv")
sp.list<-read.csv("enemy_bird_list.csv")
```

```{r,datasummary,warning=F}
std <- function(x) sd(x)/sqrt(length(x))

##clean data 
drops<-c("Grand.Total","X.blank.")##clean up df
birds=birds[,!(names(birds) %in% drops)]
birds[is.na(birds)]<-0
cols<-names(birds)
drops<-c("DIVERSITY","PLOT","VISIT")
sps<-cols[-c(1:3)]##this is a list of the names of the birds only
birds$VISIT<-as.character(birds$VISIT)
birds$total_abun<-rowSums(birds[,sps])
bird.sp<-birds[,sps]#matrix plot x sp


##summarize data for plots/analyses
#by plot
#abundance
bird.visit<-birds%>%
  mutate(birds_visit=rowSums(.[sps]))
bird.visit$VISIT<-as.character(bird.visit$VISIT)
bird.visit$PLOT<-as.character(bird.visit$PLOT)
bird.plot<-birds%>%
  mutate(birds_plot_visit=rowSums(.[sps]))%>%
  group_by(DIVERSITY,PLOT)%>%
  summarize(total_birds_plot=sum(birds_plot_visit),
            n_visit=length(VISIT),
            mean_birds_visit=mean(birds_plot_visit),
            se_birds_visit=std(birds_plot_visit))

bird.div<-bird.plot%>%
  group_by(DIVERSITY)%>%
  summarize(total_birds_div=sum(total_birds_plot),
            n_plots=length(PLOT),
            mean_birds_plot_visit=mean(mean_birds_visit),
            se_birds_plot_visit=std(mean_birds_visit))

birds.melt<-melt(birds,id.vars=c("DIVERSITY","PLOT","VISIT","total_abun"),variable.name="ID")%>%
  left_join(sp.list[,c("ID","feeding.guild","Family","Order")],by="ID")
insect.birds<-birds.melt%>%
  filter(feeding.guild %in% c("IN","FL","OM"))

ins.plot<-insect.birds%>%
  group_by(DIVERSITY,PLOT,VISIT)%>%
  summarize(total_birds_plot_visit=sum(value))%>%
  group_by(DIVERSITY,PLOT)%>%
  summarize(total_birds_plot = sum(total_birds_plot_visit),
            mean_birds_visit = mean(total_birds_plot_visit),
            n_visit=as.numeric(max(VISIT)),
            se_birds_visit = std(total_birds_plot_visit))
```
#Bird Observations  
###All Birds  
  
```{r, birdguilds,fig.height=4.5}
div.guild<-birds.melt%>%###this is the total number of birds~diversity+feeding guild
  group_by(DIVERSITY,feeding.guild)%>%
  summarize(total_birds = sum(value))

plot.visit<-birds.melt%>%
  group_by(DIVERSITY,PLOT,VISIT)%>%
  summarize(total_birds_plot=sum(value))%>%
  group_by(DIVERSITY,PLOT)%>%
  summarize(mean_birds_plot = mean(total_birds_plot))

plot.guild<-birds.melt%>%
  group_by(DIVERSITY,PLOT,VISIT,feeding.guild)%>%
  summarize(total_birds_visit = sum(value))%>%
  group_by(DIVERSITY,PLOT,feeding.guild)%>%
  summarize(mean_birds_visit=mean(total_birds_visit),se_birds_visit = std(total_birds_visit))%>%
  left_join(plot.visit[,c("PLOT","mean_birds_plot")],by="PLOT")%>%
  mutate(per_birds_visit= (mean_birds_visit/mean_birds_plot)*100)

guilddiv<-ggplot(div.guild,aes(y=total_birds,x=reorder(feeding.guild,total_birds),fill=DIVERSITY))+
  geom_bar(stat="identity")+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=wes_palette("Darjeeling"))+
  theme_minimal()+theme(legend.position="top")+labs(x="Feeding Guild",y="Total Birds Observed")

abun_by_plot<-ggplot(bird.plot,aes(x=reorder(PLOT,mean_birds_visit),y=mean_birds_visit,fill=DIVERSITY))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=mean_birds_visit-se_birds_visit,ymax=mean_birds_visit+se_birds_visit))+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=wes_palette("Darjeeling"))+
  theme_minimal()+
  theme(legend.position="none")+
  labs(x="Plot",y="Birds Observed/Visit")

plot_grid(guilddiv,abun_by_plot)
```

<br>  
```{r,fig.height=4, fig.width=4}
div_abun_plot<-ggplot(bird.div,aes(x=DIVERSITY,y=mean_birds_plot_visit,color=DIVERSITY))+
  geom_point(size=10,shape=19)+
  geom_errorbar(aes(ymin=mean_birds_plot_visit-se_birds_plot_visit,ymax=mean_birds_plot_visit+se_birds_plot_visit),color="black",width=.1)+
  labs(x="Plot Diversity",y="Bird Observed/Visit")+
  theme_minimal()+
  scale_color_manual(labels=c("Monoculture","Polyculture"),values=wes_palette("Darjeeling"))+theme(legend.position="none")
div_abun_plot+geom_point(data=bird.plot,aes(x=DIVERSITY,y=mean_birds_visit),size=2,shape=1,alpha=.65)
```  
<br>  
  
Bird Abundance ~ Plot  
```{r,abunbydivd, results="asis",}
abun.by.div<-aov(birds_visit~DIVERSITY+Error(PLOT),data=bird.visit)
pander(summary(abun.by.div),round=c(0,2,2,3,3))
```
<style type="text/css">
.table {

    width: 50%;

}
</style>

ANOVA testing for main effects of tree diversity with plot as random effect.  
<br> 
PLOT is random effect, but shoudl it be? 
    
**`r length(unique(birds.melt$PLOT))` plots** visited 4x each   
**`r sum(birds.melt$value)` bird observations** of **`r length(unique(birds.melt$ID))` species** from **`r length(unique(birds.melt$Family))` families**(`r sort(unique(birds.melt$Family))`) and **`r length(unique(birds.melt$Order))` orders**(`r sort(unique(birds.melt$Order))`)  

Guilds: Granivorous (GR), Carnivorous (CA), Nectivorous (NE), Frugivorous (FR), Insectivorous (FL = flycatching, IN= gleaning), Omnivorous (OM), NA = unidentified  

###Insectivorous Birds   
(including gleaners, flycatchers, and omnivores that consume arthropods)  

```{r, insguilds,fig.height=4.5}
ins.guild<-insect.birds%>%###this is the total number of insecitvorous birds~diversity+feeding guild
  group_by(DIVERSITY,feeding.guild)%>%
  summarize(total_birds = sum(value))

guilddiv<-ggplot(ins.guild,aes(y=total_birds,x=reorder(feeding.guild,total_birds),fill=DIVERSITY))+
  geom_bar(stat="identity")+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=wes_palette("Darjeeling"))+
  theme_minimal()+theme(legend.position="top")+labs(x="Feeding Guild",y="Total Birds Observed")

abun_by_plot<-ggplot(ins.plot,aes(x=reorder(PLOT,mean_birds_visit),y=mean_birds_visit,fill=DIVERSITY))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=mean_birds_visit-se_birds_visit,ymax=mean_birds_visit+se_birds_visit))+
  scale_fill_manual("Plot Diversity",labels=c("Monoculture","Polyculture"),values=wes_palette("Darjeeling"))+
  theme_minimal()+
  theme(legend.position="none")+
  labs(x="Plot",y="Birds Observed/Visit")

plot_grid(guilddiv,abun_by_plot)
```


**`r sum(insect.birds$value)` observations** of **`r length(unique(insect.birds$ID))` species** from **`r length(unique(insect.birds$Family))` families** (`r sort(unique(insect.birds$Family))`) and **`r length(unique(insect.birds$Order))` orders**(`r sort(unique(insect.birds$Order))`)  
  
```{r,insbytotal,fig.height=3.5,fig.width=3.5}
compareins<-left_join(bird.plot,ins.plot,by="PLOT")
ins.by.total<-lm(mean_birds_visit.x~mean_birds_visit.y,data=compareins)
p<-summary(ins.by.total)$coefficients[,4]
r2<-summary(ins.by.total)$r.squared

typeplot<-ggplot(compareins,aes(x=mean_birds_visit.x,y=mean_birds_visit.y,fill=DIVERSITY.x))+
  geom_point(size=2)+geom_smooth(method="lm",se=F)+
  geom_errorbar(aes(ymin=mean_birds_visit.y-se_birds_visit.y,ymax=mean_birds_visit.y+se_birds_visit.y))+
  geom_errorbarh(aes(xmin=mean_birds_visit.x-se_birds_visit.x,xmax=mean_birds_visit.x+se_birds_visit.x))+
  theme_minimal()+
  scale_fill_manual(values=wes_palette("Darjeeling"))+
  theme(legend.position="none")+
  labs(x="Birds/Visit",y="Insectivores/Visit")
typeplot

summary(ins.by.total)
```
    
###Exploratory Analysis  
The following figures look at characteristics of the bird observation data, and the relationships between species abundance, occurance, and sampling effort.  
```{r,EDA}
spc_pres<-apply(bird.sp>0,2,sum)
par(mfrow=c(2,2))
occur<-plot(sort(spc_pres),log='y',main="Species by plot",xlab='Cumulative Count of Species',ylab='Number of Plots')#cumulative plot of species(x) by plot(y)
##cumulative distribution of species occuranes
histo<-hist(log(spc_pres),10,main="Species Occurrance Histogram",xlab="Species Occurrances (log)")#histogram

tmp<-apply(bird.sp,2,sum)
spc_mean<-tmp/spc_pres #avg cover for each species
cumsp<-plot(sort(spc_mean),main="Cumulative Species Abundance",
     xlab="Cumulative Number of Species", ylab="Mean Abundance")

##is  mean abun of sp correlated with # of plots occur in?
abunplotcorr<-plot(spc_pres,spc_mean,main="Abundance vs Occurance",xlab="Occurrance",ylab="Mean Abundance")
#yes

#is total abundance correlated with number of species?
plt_pres<-apply(bird.sp>0,1,sum)
abuncorsp<-plot(sort(plt_pres),main="Richness by effort",xlab="Plots Visited",ylab="Number of Species/Plot")
#yes

##total abun on each plot
plt_sum<-apply(bird.sp,1,sum)##calculate total abundance for each plot
abunbyrich<-plot(sort(plt_sum),main="Abundance by effort",xlab="Plots Visited",ylab="Abundance/Plot")

##relationship between number of species/plot and total abundance
spabund<-plot(plt_pres,plt_sum,main="Abundance & Richness",xlab="Number of Species/Plot",ylab="Total Abundance")##number of species/plot, total abundance
```
  
The average species is found in `r round(mean(tmp/spc_pres),2)` plots  
The mean # of species per plot is `r round(mean(plt_pres),2)`    
The mean abundance per plot is `r round(mean(plt_sum),2)`  

### Diversity Effects  
**Bird abundance ** ~ Tree Diversity   
```{r,abunbydiv, results="asis",}
abun.by.div<-aov(lm(mean_birds_visit~DIVERSITY,data=bird.plot))
pander(summary(abun.by.div),round=c(0,2,2,3,3))
```

**Feeding Guild composition**  



```{r,fig.height=4, fig.width=4.5}
div_abun_plot<-ggplot(bird.div,aes(x=DIVERSITY,y=mean_birds_plot_visit,color=DIVERSITY))+
  geom_point(size=10,shape=19)+
  geom_errorbar(aes(ymin=mean_birds_plot_visit-se_birds_plot_visit,ymax=mean_birds_plot_visit+se_birds_plot_visit),color="black",width=.1)+
  labs(x="Plot Diversity",y="Mean birds/visit")+
  theme_minimal()+
  scale_color_manual(labels=c("Monoculture","Polyculture"),values=wes_palette("Darjeeling"))+theme(legend.position="none")
div_abun_plot+geom_point(data=bird.plot,aes(x=DIVERSITY,y=mean_birds_visit),size=2,shape=1,alpha=.65)
```

  
#Predation Experiment  

#Data  
###Description
These data are from a field experiment conducted July 2015 at the [UADY forest diversity plots near Muna, Yucatan, Mexico](http://www.treedivnet.ugent.be/ExpUADY.html). Using the tree diversity manipulation we test the predictions that  
(1) tree diversity promotes the abundance and diversity of predators (insectivorous birds) and   
(2) via effects on predators, tree diversity is correlated with greater herbivore suppression.  
  
###Data sheets  
[**Bird Observations**](https://github.com/collnell/Enemies/blob/master/enemy_bird_visit.csv)  
Bird survey results from 15 minute point counts at the plot level. Each plot was visited 4 times on separate days between 0600 and 0930. These observations reflect birds actively foraging in the plots, excluding birds flying overhead or through the plot. The species of trees used by foraging birds were recorded once per individual tree. Birds that were not able to be identified definitively are recorded as species "X", or "CH X" if a partial ID was known (CH X = chipe unknown (warbler). All bird IDs match species information on the bird list.     
  
[**Bird List**](https://github.com/collnell/Enemies/blob/master/enemy_bird_list.csv)   
Inventory on all bird species recorded at the UADY diversity plots in January and July of 2015.  Contians species codes and associated information for each ID (including feeding guild).  

**Predation Experiment**  
These data are from a clay caterpillar experiment used to assay predation rates between the plots. For each experimental plot, 2 clay models were placed on opposite branches of 12 trees. Models were visited twice, 24 and 48 hours after placement, to record attack marks. Marked models were identified as either Arthropod, Bird, and Mammal and replced with a new model.  

**Experimental Design**  
Data on plots used in experiment. Plot diversity, tree species, ETC  


