---
title: "Tree Diversity Effects"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    theme: flatly
---

```{r setup, include=FALSE}
library(shiny)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(plotly)
library(dplyr)
library(flexdashboard)
###enter data
birds<-read.csv("/Users/colleennell/Documents/R/enemies/enemy_bird_visit.csv")
sp.list<-read.csv("/Users/colleennell/Documents/R/enemies/enemy_bird_list.csv")
```

Tree Diversity Effects {.sidebar}
==============================================

**Does tree diversity mediate top-down effects of insectivorous birds?**  
  
These data reflect 10-minute bird point counts among plots of either 1 tree species (monoculture) or 4 tree species (polyculture). In addition, clay caterpillars were placed in tree canopies to assay predation rates by birds. 
  
Use the following options to select the response variable of interest and explore the results.   


```{r,datasummary}
std <- function(x) sd(x)/sqrt(length(x))

##clean data 
drops<-c("Grand.Total","X.blank.")##clean up df
birds=birds[,!(names(birds) %in% drops)]
birds[is.na(birds)]<-0
cols<-names(birds)
drops<-c("DIVERSITY","PLOT","VISIT")
sps<-cols[-c(1:3)]##this is a list of the names of the birds only
birds$VISIT<-as.character(birds$VISIT)
birds$total_abun<-rowSums(birds[,sps])
##summarize data for plots/analyses
#by plot
#abundance
bird.plot<-birds%>%
  mutate(birds_plot_visit=rowSums(.[sps]))%>%
  group_by(DIVERSITY,PLOT)%>%
  summarize(total_birds_plot=sum(birds_plot_visit),
            n_visit=length(VISIT),
            mean_birds_visit=mean(birds_plot_visit),
            se_birds_visit=std(birds_plot_visit),
            birds_30=mean_birds_visit*3,
            se_30=std(birds_plot_visit)*3)



#by diversity
abun_by_plot<-ggplot(bird.plot,aes(x=reorder(PLOT,birds_30),y=birds_30,fill=DIVERSITY))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=birds_30-se_30,ymax=birds_30+se_30))+
  labs(x="Plot",y="Birds/30/plot")


```

```{r}
#Define inputs
selectInput('bird_metric',label='Bird Community Metric',choices=c('abundance','diversity','evennes','richness'))

response<-if (input$bird_metric==abundance){
  r<-mean_birds_visit
}else{
  r<-birds_30
}





```

Bird Observations  
==========================================  


Row 
-----------------------------------------------------------------------

### Diversity Effects  
  

```{r}

abun_by_plot<-ggplot(bird.plot,aes(x=reorder(PLOT,birds_30),y=birds_30),fill=DIVERSITY))+
  geom_bar(stat="identity")+
  labs(x="Plot",y="Birds/30/plot")
abun_by_plot

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}

```

### Chart C

```{r}

```

